Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-18 docs/proposals/NSF/ProjDesc/ProjDesc.tex
NM

Upon developing and improving numerical methods to simulate mechanical
and chemical erosion, we plan to develop methods to simulate transport,
gravitational sedimentation, and external buoyant forces of the eroding grains. In addition to the new previously described quadrature, preconditioning, adaptivity, and fast summation methods,  this extension also requires contact forces between bodies to eliminate unphysical contact that is inevitable due to discretization errors. Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research.  One approach is to use introduce a potential with short-range repulsion, but such methods often introduce stiffness. Alternatively, a space-time interference volume can guarantee a minimum separation between grains.  In a fluid context, this method was first introduced by Lu et al.~\cite{lu-rah-zor2017}, and extended to allow for much smaller separation distances of rigid bodies by the PI's previous graduate student~\cite{bys-sha-qua2019}. One PhD student, who will be jointly supervised by both PIs, will simulate sedimentation of eroding bodies by developing algorithms that resolve quadrature error and contact.

@@

More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties and the flow.  The output of the high-fidelity simulations of the grain configuration can be transformed into essentially an image format, enabling us to use cutting-edge advancements on deep neural-networks for image analysis~\cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present and thus largely averts the risk of overfitting data. Additionally, high-fidelity flow conditions can be used to train a neural network, and this has been successfully used by others to predict, as one example, an anisotropy tensor~\cite{ling2016reynolds}.  Both PIs will supervise a graduate student who will design and implement the deep-neural network for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that utilize the multiphase framework to enable more efficient and larger scale coarse-grained simulations of dynamic erosion.

>>>>

Upon developing and improving numerical methods to simulate mechanical
and chemical erosion, we plan to develop methods to simulate transport,
gravitational sedimentation, and external buoyant forces of the eroding grains. In addition to the new previously described quadrature, preconditioning, adaptivity, and fast summation methods,  this extension also requires contact forces between bodies to eliminate unphysical contact that is inevitable due to discretization errors. Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research \cite{yan2019computing}.

\todo[inline]{Bryan, do you know any other references to include after `topic of current research'? I have Shelley and Yan's method. There were a bunch of others at Oaxaca but I don't remember names.}

One approach is to use introduce a potential with short-range repulsion, but such methods often introduce stiffness. Alternatively, a space-time interference volume can guarantee a minimum separation between grains.  In a fluid context, this method was first introduced by Lu et al.~\cite{lu-rah-zor2017}, and extended to allow for much smaller separation distances of rigid bodies by the PI's previous graduate student~\cite{bys-sha-qua2019}. One PhD student, who will be jointly supervised by both PIs, will simulate sedimentation of eroding bodies by developing algorithms that resolve quadrature error and contact.

@@

More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks \cite{ling2016reynolds, kutz2017deep, ling2016machine, pelt2018mixed} to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties and the flow.  The output of the high-fidelity simulations of the grain configuration can be transformed into essentially an image format, enabling us to use cutting-edge advancements on deep neural-networks for image analysis~\cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present and thus largely averts the risk of overfitting data. Additionally, high-fidelity flow conditions can be used to train a neural network, and this has been successfully used by others to predict, as one example, an anisotropy tensor~\cite{ling2016reynolds}.  Both PIs will supervise a graduate student who will design and implement the deep-neural network for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that utilize the multiphase framework to enable more efficient and larger scale coarse-grained simulations of dynamic erosion.
